Role of the microstructure in the failure of a protein gel under stress

### 2014-10-13 poster_Palavas.tex
Mise en page avec minipage plutôt que TextPos, garde la grille

@@

@@

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
\def\Norulehead#1{\noindent\hbox to \hsize{\hfil\LARGE\textcolor{Main}{\textsf{#1}}}\bigskip}
\def\Head#1{\Norulehead{\dotfill #1}}
\def\LHead#1{{\LARGE #1}\smallskip}
\def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{\uppercase{#1}}}}

@@

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
%\begin{textblock}{15}(0,0)
\begin{center}
\Title{
Role of the microstructure in the failure of a protein gel under stress
}
\end{center}

\smallskip
\LHead{Mathieu Leocmach, Christophe Perge, Thomas Gibaud, Sébastien Manneville
}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}

@@

%\end{textblock}

\begin{textblock}{11}(0,3)
	\Head{Over-acidified protein gels: beyond isoelectric pH}

	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})

	\tikzsetnextfilename{prise_cas4}
	\input{fig_poster_palavas/prise_cas4.pgf}

\end{textblock}

\begin{textblock}{11}(0,8.5)
	\Head{Linear rheology, from power-law to glassy}
	\tikzsetnextfilename{sweep_cas4}
	\input{fig_poster_palavas/sweep_cas4.pgf}

\end{textblock}

\begin{textblock}{3}(12,8.5)
	\Head{Microscopy}
	Final state without over-acidification

@@

\end{textblock}

\begin{textblock}{11}(0,11.5)
	\Head{Evolution of the microscructure: preserved network, swelling aggregates}
	\tikzsetnextfilename{structure_confocal}
	\input{fig_poster_palavas/structure_confocal.pgf}

\end{textblock}

\begin{textblock}{11}(0,15.5)
	\Head{Three creep regimes: much more chaotic when over-acidified}
	\tikzsetnextfilename{creep_cas4_gdl4}
	\input{fig_poster_palavas/creep_cas4_gdl4.pgf}

\end{textblock}

\begin{textblock}{3}(12,15.5)
	\Head{Monkman-Grant}
	Predict failure from the minimum

	\tikzsetnextfilename{MonkmanGrant}
	\input{fig_poster_palavas/MonkmanGrant.pgf}

\end{textblock}

\begin{textblock}{11}(0,18.5)
	\Head{Basquin law rescaling}
	\begin{tabu}{XX}
	\Subhead{Composition} rescaled by elastic modulus&
	\Subhead{Geometry} rescaled as a fracture nucleation rate
	\end{tabu}

	\tikzsetnextfilename{basquin}
	\input{fig_poster_palavas/basquin.pgf}

\end{textblock}

\begin{textblock}{3}(12,20)
	\Head{Wavelength}
	\tikzsetnextfilename{lambda}
	\input{fig_poster_palavas/lambda.pgf}

\end{textblock}

\begin{textblock}{15}(0,24.75)
\small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
\end{textblock}

\textblockcolour{lightgray!50!white}
\TPMargin*{ 0.125\TPHorizModule }
\begin{textblock}{2.875}(12,3.125)%real width of 3
	\Norulehead{Take home}
	Spontaneous wrinkling of over-acidified casein gels in non-adhesive confinement

	\Subhead{Original patterns}
	\begin{itemize}
		\item nested (optical elements ?)
		\item kinetic wavelength selection
	\end{itemize}
	\Subhead{A new wrinkling mechanism}
	\begin{itemize}
		\item porous flow resist bending
		\item Darcy can beat Poiseuille
		\item may extend to other porous sheets\linebreak  (wet paper, biological membranes\ldots)
	\end{itemize}
	\Subhead{Viscosity influences}
	\begin{itemize}
		\item porous dissipation
		\item gel formation\hfill
		\raisebox{0.6\normalbaselineskip}[0pt][0pt]{$\left.\rule{0pt}{1.1\normalbaselineskip}\right\}$ thus wrinkling}
	\end{itemize}

\end{textblock}%Conclusion

>>>>

\usepackage[margin=40mm]{geometry}

@@

\usepackage{framed}
\colorlet{shadecolor}{lightgray!50!white}

@@

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
\def\Norulehead#1{\noindent\hbox to \hsize{\hfil\LARGE\textcolor{Main}{\textsf{#1}}}\bigskip}
\def\Head#1{\Norulehead{\dotfill #1}}
\def\LHead#1{{\LARGE #1}\smallskip}
\def\Subhead#1{\noindent{\large\color{Accent1}\textsc{#1}}}
\def\Title#1{\noindent{\VeryHuge\color{Accent2}\raggedright\textsf{\fontspec[BoldFont=Lovelo-LineBold]{Lovelo-LineBold}{#1}}}}

@@

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.

\begin{center}
\Title{
Role of the microstructure in the failure of a protein gel under stress
}
\end{center}

\vfill
\LHead{Mathieu Leocmach, Christophe Perge, Thomas Gibaud, Sébastien Manneville
}\hfill\texttt{\color{Accent1}mathieu.leocmach@ens-lyon.fr}

@@

\vfill
\begin{minipage}[b][21.5\TPVertModule]{11\TPHorizModule}
	\Head{Over-acidified protein gels: beyond isoelectric pH}

	Milk protein solution (4\%w sodium caseinate) acidified by the slow hydrolysis of glucono-$\delta$-lactone (\textsc{gdl})\\ \textit{Bremer et al., Colloids Surf. (1990); van Vliet \& Walstra, Faraday Discuss. (1995); Lucey \& Singh, Food Res. Int. (1997); Leocmach et al. PRL (2014)}

	\smallskip
	\tikzsetnextfilename{prise_cas4}
	\input{fig_poster_palavas/prise_cas4.pgf}
	\begin{tabu}{X[5]XX[5]}
	Caseins regain solubility below isoelectric $\Rightarrow$ overshoot in $G^\prime$&&
	Single time scale $(t_m-t_i)$ governs gelation and softening.\\
	How does it translate in term of microstructure?&&Diverge when $pH_\infty\rightarrow pI$
	\end{tabu}

	\vfill
	\Head{Linear rheology, from power-law to soft glassy}
	\tikzsetnextfilename{sweep_cas4}
	\input{fig_poster_palavas/sweep_cas4.pgf}
	\begin{tabu}{X[4]X[2]X[5]}
	Near isoeletric, power law rheology typical of gels.&&
	Minimum in loss modulus appear with over-acidification.\\
	Flatter with over-acidification.&%&
	\multicolumn{2}{l}{\textcolor{Accent1}{$\blacktriangleright$} Over-acidified behaviour typical of soft glassy materials}\\
	&\textit{Sollich, PRE (1998)}&\textit{Mason, Curr. Opin. Colloid Interface Sci. (1999)}\\
	\end{tabu}

	\vfill
	\Head{Evolution of the microscructure: preserved network, swelling aggregates}
	\tikzsetnextfilename{structure_confocal}
	\input{fig_poster_palavas/structure_confocal.pgf}
	\begin{tabu}{X[5]XX[5]}
	Maximum $G^\prime$ corresponds to maximum contrast between phases $\chi$&&
	\textcolor{Accent1}{$\blacktriangleright$} Strongest gel when attraction is strongest\\
	Characteristic size $\xi$ increases during over-acidification&&
	\textcolor{Accent1}{$\blacktriangleright$} Casein aggregates swell $\sim$ over compressed gel $\Rightarrow$ soft glassy\\
	\end{tabu}

	\vfill
	\Head{Three creep regimes: much more erratic when over-acidified}
	\tikzsetnextfilename{creep_cas4_gdl4}
	\input{fig_poster_palavas/creep_cas4_gdl4.pgf}
	\begin{tabu}{X[4]X[3]X[3]}
	$\tau_f$: divergence of the strain $\gamma$, failure time&
	\multicolumn{2}{l}{Robust Andrade creep, compatible with power-law rheology \hfill\textit{Leocmach et. al, PRL (2014)}}\\
	$\tau_\text{min}$: minimum of shear rate $\dot{\gamma}$, fracture nucleation&
	Robust finite time divergence&
	\textcolor{Accent1}{$\blacktriangleright$}Possible to reform bonds? Glassy?
	\end{tabu}

	\vfill
	\Head{Predict fracture nucleation time: Basquin law rescaling}
	\begin{tabu}{XX}
	\hspace{\TPHorizModule}\Subhead{Composition} rescaled by elastic modulus&
	\hspace{\TPHorizModule}\Subhead{Geometry} rescaled as a fracture nucleation rate
	\end{tabu}

	\tikzsetnextfilename{basquin}
	\input{fig_poster_palavas/basquin.pgf}
	\begin{tabu}{X[5]XX[5]}
	Robust Basquin law despite glassy behaviour&&\textcolor{Accent1}{$\blacktriangleright$} Failure determined by network breaking \& fracture nucleation
	\end{tabu}

\end{minipage}
\hfill
\begin{minipage}[b][21.5\TPVertModule]{3\TPHorizModule}
\begin{shaded*}
	\Norulehead{Take home}
	\Subhead{Over acidification}
	\begin{itemize}
		\item protein aggregates swell
		\item power-law rheology, decreasing exponent
		\item dissipation becomes glassy
	\end{itemize}
	\Subhead{Creep response}
	\begin{itemize}
		\item Andrade creep
		\item finite time singularity
		\item erratic transition
	\end{itemize}
	\Subhead{Fracture time prediction}
	\begin{itemize}
		\item composition influence limited to $G^\prime$
		\item geometry factors as a nucleation rate
	\end{itemize}
\end{shaded*}

	\vfill
	\Head{Microscopy}
	Final state without over-acidification\\

@@

	\vfill
	\Head{Monkman-Grant}
	Predict failure from the minimum

	\tikzsetnextfilename{MonkmanGrant}
	\input{fig_poster_palavas/MonkmanGrant.pgf}
	Height-dependent near isoelectric\\
	Upper bound for over-acidified

	\vfill
	\Head{Wavelength}
	\tikzsetnextfilename{lambda}
	\input{fig_poster_palavas/lambda.pgf}
	Compatible with nucleation rate

\end{minipage}

\vfill
\small{\texttt{The research leading to these results has received funding from the European Research Council under the European Union's Seventh Framework Programme (FP7/2007-2013) / ERC grant agreement n°~258803.}}
